---
title: "20211102 Logistic regressions (Uni students)"
author: "cheungngo"
date: "2 November 2021"
output: html_document
---

### Loading data

```{r}
library(readxl)
uni_students <- read_excel("~/unrest_covid_mental/data/uni_students.xlsx")
```

### Function required

##### Serial logistic regression for continuous variables

```{r}
log_reg_spss_cont0 = function(datay, datax, j, i) {
  var = as.numeric(unlist(datax[,i]))
  dep = as.numeric(unlist(datay[,j]))
  ind = which(!is.na(var))
  var = var[ind]
  dep = dep[ind]
  coeff = matrix(ncol = 4)
  modglm = glm(dep~var, family = "binomial")
  coeff[1,] = summary(modglm)$coef[2,]
  rownames(coeff) = colnames(datax)[i]
  colnames(coeff) = c("Estimate", "SD", "z-value", "p-value")
  coeff = round(coeff, digits = 3)
  return(coeff)
}

cont_beau = function(cont) {
  OR = apply(cont, 1, function (i) {
    exp(i[1] + qnorm(c(0.025,0.5,0.975)) * i[2])
  })
  OR = t(OR)
  OR_b = apply(OR, 1, function (i) {
    if (i[2] != 1) {
      return(paste(round(i[2],2), " (", round(i[1],2), "-", round(i[3],2), ")", sep = ""))
    } else {
      return(1)
    }
  })
  result = cbind(OR_b, cont[,4])
  colnames(result) = c("OR","p")
  return(result)
}

log_reg_cont_care4all = function (datay, datax, j, i) {
  for (x in i) {
    assign(paste("mod", colnames(datax)[x],sep = ""),
           log_reg_spss_cont0(datay, datax, j, x))
  }
  tab = eval(parse(text = paste("mod", colnames(datax)[i[1]],sep = "")))
  for (x in i[-1]) {
    tab2 = eval(parse(text = paste("mod", colnames(datax)[x],sep = "")))
    tab = rbind(tab, tab2)
  }
  return(cont_beau(tab))
}
```


##### Serial logistic regression for categorical variables

```{r}
OR_CI = function(m) {
  ifelse(is.na(coef(m)[2]),
         return(c(NA,NA,NA)),
         return(exp(summary(m)$coefficients[2,1] +
                      qnorm(c(0.025,0.5,0.975)) * summary(m)$coefficients[2,2])))
}

log_reg_spss_kref_care = function(datay, datax, j, i, kref) {
  var = as.numeric(unlist(datax[,i]))
  dep = as.numeric(unlist(datay[,j]))
  ind = which(!is.na(var))
  var = var[ind]
  dep = dep[ind]
  coeff = matrix(nrow = length(table(var)),
                 ncol = 4)
  coeff[1,] = c(NA,1,NA,NA)
  ref = kref
  level = as.numeric(unlist(levels(as.factor(var))))
  level = level[-which(level==ref)]
  count = 2
  for (k in level) {
    ind2 = which(var==k | var==ref)
    yy = dep[ind2]
    xx = var[ind2]
    modglm = glm(yy~xx, family = "binomial")
    coeff[count,1:3] = OR_CI(modglm)
    ifelse(is.na(coef(modglm)[2]),
           coeff[count,4] <- NA,
           coeff[count,4] <- as.numeric(unlist(summary(modglm)$coefficients[2,4])))
    count = count + 1
  }
  
  var_names = colnames(datax)[i]
  rownames(coeff) = paste(var_names[1], c(ref, level), sep = "_")
  colnames(coeff) = c("LCI", "OR", "UCI", "p-value")
  coeff = round(coeff, digits = 3)
  return(coeff)
}

beautify_cate = function (data) {
  data[,1:3] = apply(data[,1:3], 2, function(i) {
    round(i, digits = 1)
  })
  OR_b = apply(data, 1, function (i) {
    if (i[2] != 1) {
      return(paste(i[2], " (", i[1], "-", i[3], ")", sep = ""))
    } else {
      return(1)
    }
  })
  result = cbind(OR_b, data[,4])
  colnames(result) = c("OR","p")
  return(result)
}

log_reg_cate_care4all = function (datay, datax, j, i, kref) {
  for (x in i) {
    assign(paste("mod", colnames(datax)[x],sep = ""),
           log_reg_spss_kref_care(datay, datax, j, x, kref))
  }
  tab = eval(parse(text = paste("mod", colnames(datax)[i[1]],sep = "")))
  for (x in i[-1]) {
    tab2 = eval(parse(text = paste("mod", colnames(datax)[x],sep = "")))
    tab = rbind(tab, tab2)
  }
  return(beautify_cate(tab))
}
```

##### A function to create ordinal variables from multiple columns of boolean variables

```{r}
ordinal_from_bools = function(data, cols_as_list, start) {
  s = start
  ord = rep(start, dim(data)[1])
  for (i in cols_as_list) {
    s = s + 1
    for (j in i) {
      x = which(unlist(data[,j])==1)
      ord[x] = s
    }
  }
  return(ord)
}
```

##### Functions to create a series of confusion tables

```{r}
tab_ngo_0 = function(datay, datax, y, x) {
  datax_n = datax
  datay_n = datay
  tab = table(unlist(datax_n[,x]),
              unlist(datay_n[,y]))
  rownames(tab) = paste(colnames(datax)[x],
                        rownames(tab),
                        sep = "_")
  tab_new = apply(tab, 2, function (i) {
    sapply(i, function (j) {
      paste(j, " (", round(j / sum(i) * 100, digits = 2), "%)", sep = "")
    })
  })
  return(tab_new)
}

serial_tab_0 = function(datay, datax, y, spec_x) {
  for (i in 1:length(spec_x)) {
    assign(paste("tab", colnames(datax)[spec_x[i]],sep = ""),
           tab_ngo_0(datay, datax, y, spec_x[i]))
  }
  tab = eval(parse(text = paste("tab", colnames(datax)[spec_x[1]],sep = "")))
  for (i in 2:length(spec_x)) {
    tab2 = eval(parse(text = paste("tab", colnames(datax)[spec_x[i]],sep = "")))
    tab = rbind(tab, tab2)
  }
  return(tab)
}
```

##### Function for a series of distributions

```{r}
cont_dist_ngo = function (datay, datax, y, x) {
  ind = which(datay[,y] == 0)
  mat = matrix(nrow = 1,
               ncol = 2)
  datax = unlist(datax[,x])
  x1 = datax[ind]
  x2 = datax[-ind]
  ifelse((length(unique(x1))>2 | length(unique(x2))>2),
         {
           mn1 = round(mean(x1), digits = 2)
           mn2 = round(mean(x2), digits = 2)
           sd1 = round(sd(x1), digits = 2)
           sd2 = round(sd(x2), digits = 2)
           mat[1,1] <- paste(mn1, " (", sd1, ") ", sep = "")
           mat[1,2] <- paste(mn2, " (", sd2, ") ", sep = "")
         },
         mat[1,1:2] <- c(NA,NA))
  return(mat)
}

serial_cont_dist = function(datay, datax, y, spec_x) {
  mat = matrix(nrow = length(spec_x),
               ncol = 2)
  for (i in 1:length(spec_x)) {
    mat[i,] = cont_dist_ngo(datay, datax, y, spec_x[i])
  }
  colnames(mat) = c("mean(SD) (y=0)", "mean(SD) (y=1)")
  rownames(mat) = colnames(datax)[spec_x]
  return(mat)
}

```


##### A function to process all logistic regression at one time

```{r}
serial_log_reg = function(datay, datax, target_column, ref) {
  
  t = target_column
  
  tab = matrix(c("OR", "p"), byrow = T, nrow = 1, ncol = 2)
  rownames(tab) = "Var"
  
  p01 = c()
  
  for (i in 1:dim(uni_new)[2]) {
    if ( (ref[i] == "cont") | (ref[i] == "cont_ord") ) {
      tab_new = log_reg_cont_care4all(datay, datax, t, i)
      for (p in tab_new[,2]) {
        if (p < 0.1) {
            p01 = c(p01, i)
            break
          }
      }
      tab = rbind(tab, tab_new)
    } else {
      tab_new = log_reg_cate_care4all(datay, datax, t, i, ref[i])
      for (p in tab_new[,2]) {
        if (!is.na(p)){
          if (p < 0.1) {
            p01 = c(p01, i)
            break
          }
        }
      }
      tab = rbind(tab, tab_new)
    }
  }
  
  return(list(as.data.frame(tab), p01))
}
```

### Function to create dummy variables for multivariate logistic regression (for ordinal variables with more than two categories)

```{r, echo=FALSE}
# A function to create dummy variables

create_dummy = function(masterdata,ind) {
  
  coll = unlist(masterdata[,ind])
  
  n = length(unique(coll))
  len = length(coll)
  
  data = matrix(nrow = len,
                ncol = n)
  count = 1
  for (i in unique(coll)) {
    data[,count] = sapply(coll, function(x) {
      ifelse(x==i, 1,0)
      })
    count = count+1
  }
  c_name = colnames(masterdata)[ind]
  colnames(data) = sapply(unique(coll), function (i) {
    paste(c_name, i, sep = "_")
  })
  return(data)
}

# To be used with create_dummy()
cbind_dummy = function(masterdata, ind) {
  ini = ind[1]
  initial = create_dummy(masterdata, ini)
  if (length(ind) >=2) {
    for (i in ind[-1]) {
      subsequent = create_dummy(masterdata, i)
      initial = cbind(initial, subsequent)
    }
    return(initial)
  } else {
    return(initial)
  }
}
```

##### Function to perform multivariate logistic regression

```{r}
library(MASS)

multivar_log_reg = function(datay, datax, ref, j, sigvar_col, dir) {
  
  sigvar = sigvar_col
  fordum = c()
  
  for (i in sigvar) {
    if ( (ref[i] != "cont") & (ref[i] != "cont_ord") ) {
      if (length(unique(datax[,i]))>2) {
        fordum = c(fordum, i)
      }
    }
  }
    
  nodum = setdiff(sigvar, fordum)
  
  if (length(fordum) > 0) {
    y_x = cbind(datax[,nodum],
                cbind_dummy(datax, fordum))
  } else {
    y_x = datax[,nodum]
  }
  
  y_x$y = datay[,j]
  
  full_model = glm(y ~.,
                   data = y_x,
                   family = "binomial")
  
  step_model = stepAIC(full_model,
                       direction = dir,
                       trace = 0)
  return(step_model)
}
```

##### Function to extract the coefficient table from the model, and beautify

```{r}
model_beau = function(model) {
  
  out = as.data.frame(cont_beau(summary(model)$coefficients))
  out$sig = sapply(out$p, function(i) {
    pval = as.numeric(i)
    if (pval < 0.001) {
      "***"
    } else if (pval < 0.01) {
        "**"
    } else if (pval < 0.05) {
        "*"
    } else if (pval < 0.1) {
        "."
    } else {
        ""
      }
  })
  return(out)
}
```

### Data Wrangling

```{r}
names(uni_students)
```

```{r}
# outcomes
uni_new_y = data.frame(uni_students[,c(126,127,130)]) # 0/1
```

```{r}
# demographics

# Gender and Age
uni_new = uni_students[,(52:53)]
# 0:F/1:M 
# age as ordinal, could be analysed as cont (i.e. no ref)
# or may use 2.2 as ref
ref = c(0,"cont_ord")

# Degree
uni_new = cbind(uni_new, uni_students[,54])
# 40 as ref, most people, bachlor
ref = c(ref,40)
### For discussion: may combine below-bachelors

# Year
# just continuous var
uni_new = cbind(uni_new, uni_students[,56])
ref = c(ref, "cont_ord")

# District
uni_new = cbind(uni_new, uni_students[,57])
# use 16 (shatin) as ref (most people)
ref = c(ref, 16)

# Psychiatric History
uni_new = cbind(uni_new, uni_students[,90])
ref = c(ref, 0)
```

```{r}
# Part C: Covid-status

# need simplification
# 0 for no contact / effects at all
# 1 for more distant effects
# 2 for close contacts
# 3 for infected cases

cs = ordinal_from_bools(uni_students, list(63:65, 59:61, 58), 0)
uni_new$C_1 = cs
ref = c(ref, 0)
```

```{r}
# Part D: Effects on daily lives by COVID

# Question 1a and 1b will be regrouped

# occupational related factors
uni_new$D_1_occu = ordinal_from_bools(uni_students, list(67:69), 0)
ref = c(ref, 0)

# Living environment / family related
uni_new$D_1_living = ordinal_from_bools(uni_students, list(70:71), 0)
ref = c(ref, 0)

#  Medical related
uni_new$D_1_medical = ordinal_from_bools(uni_students, list(72), 0)
ref = c(ref, 0)

# School related
uni_new$D_1_school = ordinal_from_bools(uni_students, list(73:75), 0)
ref = c(ref, 0)

# Psychologically related CANCELED for perfect correlation

# Q2-5 just copied to the new dataframe
uni_new = cbind(uni_new, uni_students[,85:88])
ref = c(ref, rep("cont", 4))

# Q6 CANCELLED for multicollinearity
```

```{r}
# Part E: perception about COVID
# just copy will do

uni_new = cbind(uni_new, uni_students[,91:92])
ref = c(ref, rep("cont", 2))
```

```{r}
# Part F: social unrest

# Q1 could be converted as an continuous / ordinal variable, with 0 as no contact

uni_new$F_1 = ordinal_from_bools(uni_students, list(97,96,95,94,93), 0)
# should be as "cont_ord"
ref = c(ref, "cont_ord")

# Q2 could just be copied

uni_new = cbind(uni_new, uni_students[,99])
ref = c(ref, "cont")

# Q3 should be deleted for possible multicollinearity (with Q4)

# Q4: bi 1-4 could be just copied
uni_new = cbind(uni_new, uni_students[,101:104])
ref = c(ref, rep(0,4))
# bi 5:6 as occupational related
uni_new$F_3_b_i_Occu = ordinal_from_bools(uni_students, list(105:106), 0)
ref = c(ref, 0)
# bii as one
uni_new$F_3_b_ii_Phy = ordinal_from_bools(uni_students, list(c(110,112)), 0)
ref = c(ref, 0)
# biii should be cancelled for perfect correlation
```

```{r}
uni_new = as.data.frame(uni_new)
uni_new_y = as.data.frame(uni_new_y)
```

### Distribution tables

```{r}
ref_cate = which(ref != "cont")
ref_cont = which(ref == "cont")

phq_cate = serial_tab_0(uni_new_y, uni_new, 1, ref_cate)
gad_cate = serial_tab_0(uni_new_y, uni_new, 2, ref_cate)
iesr_cate = serial_tab_0(uni_new_y, uni_new, 3, ref_cate)

phq_cont = serial_cont_dist(uni_new_y, uni_new, 1, ref_cont)
gad_cont = serial_cont_dist(uni_new_y, uni_new, 2, ref_cont)
iesr_cont = serial_cont_dist(uni_new_y, uni_new, 3, ref_cont)

phq_tabs = rbind(phq_cate, phq_cont)
gad_tabs = rbind(gad_cate, gad_cont)
iesr_tabs = rbind(iesr_cate, iesr_cont)

write.csv(phq_tabs, "~/unrest_covid_mental/data/uni_std/phq_tabs.csv")
write.csv(gad_tabs, "~/unrest_covid_mental/data/uni_std/gad_tabs.csv")
write.csv(iesr_tabs, "~/unrest_covid_mental/data/uni_std/iesr_tabs.csv")
```

### A series of univariate logistic regressions

```{r, warning=FALSE}
phqall = serial_log_reg(uni_new_y, uni_new, target_column = 1, ref = ref)
gadall = serial_log_reg(uni_new_y, uni_new, target_column = 2, ref = ref)
iesrall = serial_log_reg(uni_new_y, uni_new, target_column = 3, ref = ref)
```

```{r}
write.csv(phqall[[1]], "~/unrest_covid_mental/data/uni_std/phq_logr.csv")
write.csv(gadall[[1]], "~/unrest_covid_mental/data/uni_std/gad_logr.csv")
write.csv(iesrall[[1]], "~/unrest_covid_mental/data/uni_std/iesr_logr.csv")
```

### Multivariate logistic regression

```{r}
phq_model = multivar_log_reg(uni_new_y, uni_new, ref, 1, phqall[[2]], "both")
phq_model_b = model_beau(phq_model)
phq_model_b
write.csv(phq_model_b, "~/unrest_covid_mental/data/uni_std/phq_model_b.csv")
```


```{r}
gad_model = multivar_log_reg(uni_new_y, uni_new, ref, 2, gadall[[2]], "both")
gad_model_b = model_beau(gad_model)
gad_model_b
write.csv(gad_model_b, "~/unrest_covid_mental/data/uni_std/gad_model_b.csv")
```

```{r}
iesr_model = multivar_log_reg(uni_new_y, uni_new, ref, 3, iesrall[[2]], "both")
iesr_model_b = model_beau(iesr_model)
iesr_model_b
write.csv(iesr_model_b, "~/unrest_covid_mental/data/uni_std/iesr_model_b.csv")
```